---
title: "County Health Rankings"
output: html_notebook
---

This is an analysis of the Robert Wood Johnson Foundation County Health Rankings.

Setting the working directory
```{r set wd}
setwd("C:/Users/natek/Documents/County Health Rankings Analysis")
```

Loading libraries
```{r libraries, message = FALSE}
library(tidyverse)
```

I tried reading in the 2014, 2015, 2016, and 2017 data. Unfortunately, all of them are messy datasets and will need to be cleaned up individually before combining any of them. So I'll start with 2017 and return to the others later. 

There are two columns that caused difficulty for 'read_csv()`, so I read them in as characters and will fix them if they're required for the analysis  
```{r data}
chr_df_17 = read_csv("data/2017CHR_CSV_Analytic_Data.csv",
                       col_types = cols(
                       measure_127_value = col_character(),
                       measure_131_value = col_character()
                       )
                      )

chr_df_17 = chr_df_17 %>% rename(FIPS = fipscode) #renaming a variable for use in the function below

```

Next I'll take just our peer cities and add columns indicating baseline or current peer

```{r}

pull_all_peers <- function(data){
  data = data %>% filter(data$FIPS == 1073 |data$FIPS == 37119
                       |data$FIPS == 39061 |data$FIPS == 39049
                       |data$FIPS == 26081 |data$FIPS == 37081
                       |data$FIPS == 45045 |data$FIPS == 18097
                       |data$FIPS == 29095 |data$FIPS == 47093
                       |data$FIPS == 21111 |data$FIPS == 47157
                       |data$FIPS == 47037 |data$FIPS == 40109
                       |data$FIPS == 31055 |data$FIPS == 29189
                       |data$FIPS == 29510
                       |data$FIPS == 40143 |data$FIPS == 39113
                       |data$FIPS == 12031 |data$FIPS == 37183
                       |data$FIPS == 37183 |data$FIPS == 51760)
  
    data$baseline <- 1
    data$current <- 1
    data$baseline[data$FIPS==26081|data$FIPS==29189
                       |data$FIPS==29510|data$FIPS==40109
                       |data$FIPS==40143|data$FIPS==45045
                       |data$FIPS==47093] <- 0
    data$current[data$FIPS== 12031|data$FIPS==37183|
                        data$FIPS==39113|data$FIPS==51760] <- 0
  data
}

chr_df_17 = pull_all_peers(chr_df_17)

```

Now renaming everything to have useful names

```{r}




```

